Precise predictions for same-sign W-boson scattering at the LHC

### 2018-03-08 Draft/contributions.tex

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
For the distribution in the di-jet invariant mass (left), one can observe that the EW contribution peaks around an invariant mass of about $80\GeV$.
This is due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculations relying on the VBS approximation.
The EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$.
The same holds true for jet rapidity difference larger than $2.5$ (right).
This justifies why cuts on these two observables are used in order to enhance the EW contribution over the QCD one.
In particular, in order to have a large EW contribution, rather exclusive cuts are required.

>>>>

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distributions in the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
For the di-jet invariant-mass  distribution (left), one can observe that the EW contribution peaks around an invariant mass of about $80\GeV$.
This is due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculations relying on the VBS approximation.
The EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$.
The same holds true for jet-rapidity difference larger than $2.5$ (right).
This justifies why cuts on these two observables are used in order to enhance the EW contribution over the QCD one.
In particular, in order to have a large EW contribution, rather exclusive cuts are required.

### 2018-03-08 Draft/description.tex

  \item {\sc MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} (henceforth {\sc MG5\_aMC}) is an automatic meta-code (a code that generates codes) which makes it possible to simulate any scattering process
      including NLO QCD corrections both at fixed order and including matching to parton showers, using the {\sc MC@NLO}\ method~\cite{Frixione:2002ik}. It makes use of the FKS subtraction method~\cite{Frixione:1995ms,
        Frixione:1997np} (automated in the module {\sc MadFKS}~\cite{Frederix:2009yq,
        Frederix:2016rdc}) for regulating IR singularities. The computations of one-loop amplitudes are carried out by switching dynamically between
        two integral-reduction techniques, OPP~\cite{Ossola:2006us} or Laurent-series expansion~\cite{Mastrolia:2012bu},
        and tensor-integral reduction~\cite{Passarino:1978jh,Davydychev:1991va,Denner:2005nn}. These have been automated in the module {\sc MadLoop}~\cite{Hirschi:2011pa}, which
        in turn exploits {\sc CutTools}~\cite{Ossola:2007ax}, {\sc Ninja}~\cite{Peraro:2014cba,
        Hirschi:2016mdz}, {\sc IREGI}~\cite{ShaoIREGI}, or {\sc Collier}~\cite{Denner:2016kdg}, together with an in-house 
        implementation of the {\sc OpenLoops} optimisation~\cite{Cascioli:2011va}. Finally, scale and PDF uncertainties can be obtained in an exact manner via reweighting
        at zero additional CPU cost~\cite{Frederix:2011ss}.\\
        The simulation of VBS at NLO-QCD accuracy can be performed by issuing the following commands in the program interface:
\begin{verbatim}
> set complex_mass_scheme
> import model loop_qcd_qed_sm_Gmu
> generate p p > e+ ve mu+ vm j j QCD=0 [QCD]
> output
\end{verbatim}
  With these commands the complex-mass scheme is turned on, then the NLO-capable model is loaded\footnote{Despite
            the {\tt loop\_qcd\_qed\_sm\_Gmu} model also includes NLO counterterms for computing EW corrections, it is not yet possible to compute such corrections
        with the current public version of the code.}, finally the process code is generated (note the {\tt QCD=0} syntax to select the purely-EW process)
        and written to disk. No approximation is performed for the Born and real-emission matrix elements. 
        For what concerns the virtual matrix element, because of some internal limitations which will be lifted in the future version capable of computing both QCD and EW corrections,
        only loops with QCD-interacting particles are generated. Such an approximation is equivalent to the assumption that the finite part of
        those loops which feature EW bosons is zero. In practice, since a part of the contribution to the single pole is also missing, the internal 
        pole-cancellation check at run time has to be turned off, by setting the value of the {\tt IR\-Pole\-Check\-Threshold} and 
        {\tt Precision\-Virtual\-At\-Run\-Time} parameters in the {\tt Cards\-/FKS\_\-params.dat} file to -1.

>>>>

  \item {\sc MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} (henceforth {\sc MG5\_aMC}) is an automatic meta-code (a code that generates codes) which makes it possible to simulate any scattering process
      including NLO QCD corrections both at fixed order and including matching to parton showers, using the {\sc MC@NLO}\ method~\cite{Frixione:2002ik}. It makes use of the subtraction method by Frixione-Kunszt and Signer (FKS)~\cite{Frixione:1995ms,
        Frixione:1997np} (automated in the module {\sc MadFKS}~\cite{Frederix:2009yq,
        Frederix:2016rdc}) for regulating IR singularities. The computations of one-loop amplitudes are carried out by switching dynamically between
        two integral-reduction techniques, OPP~\cite{Ossola:2006us} or Laurent-series expansion~\cite{Mastrolia:2012bu},
        and tensor-integral reduction~\cite{Passarino:1978jh,Davydychev:1991va,Denner:2005nn}. These have been automated in the module {\sc MadLoop}~\cite{Hirschi:2011pa}, which
        in turn exploits {\sc CutTools}~\cite{Ossola:2007ax}, {\sc Ninja}~\cite{Peraro:2014cba,
        Hirschi:2016mdz}, {\sc IREGI}~\cite{ShaoIREGI}, or {\sc Collier}~\cite{Denner:2016kdg}, together with an in-house 
        implementation of the {\sc OpenLoops} optimisation~\cite{Cascioli:2011va}. Finally, scale and PDF uncertainties can be obtained in an exact manner via reweighting
        at zero additional CPU cost~\cite{Frederix:2011ss}.\\
        The simulation of VBS at NLO-QCD accuracy can be performed by issuing the following commands in the program interface:
\begin{verbatim}
> set complex_mass_scheme
> import model loop_qcd_qed_sm_Gmu
> generate p p > e+ ve mu+ vm j j QCD=0 [QCD]
> output
\end{verbatim}
  With these commands the complex-mass scheme is turned on, then the NLO-capable model is loaded\footnote{Despite
            the {\tt loop\_qcd\_qed\_sm\_Gmu} model also includes NLO counterterms for computing EW corrections, it is not yet possible to compute such corrections
        with the current public version of the code.}, finally the process code is generated (note the {\tt QCD=0} syntax to select the purely-EW process)
        and written to disk. No approximation is performed for the Born and real-emission matrix elements. 
        For what concerns the virtual matrix element, because of some internal limitations which will be lifted in the future version capable of computing both QCD and EW corrections,
        only loops with QCD-interacting particles are generated. Such an approximation is equivalent to the assumption that the finite part of
        those loops which feature EW bosons is zero. In practice, since a part of the contribution to the single pole is also missing, the internal 
        pole-cancellation check at run time has to be turned off, by setting the value of the {\tt IR\-Pole\-Check\-Threshold} and 
        {\tt Precision\-Virtual\-At\-Run\-Time} parameters in the {\tt Cards\-/FKS\_\-params.dat} file to -1.

### 2018-03-08 Draft/distributionsNLO.tex

We start with Fig.~\ref{fig:distNLO1} which displays the invariant mass (left) and the rapidity separation (right) of the two tagging jets.
For high invariant mass, all predictions agree rather well.
On the other hand, for low invariant mass, the hierarchy present at the level of the cross section is here reproduced.
The VBS-approximated predictions ({\sc Bonsay} and {\sc Powheg-Box}) are lower than the full calculation ({\sc MoCaNLO}+{\sc Recola}).
The full calculation is rather well approximated by the hybrid VBS approximation implemented in {\sc MadGraph5\_aMC\-@NLO}.
Finally, {\sc VBFNLO} which includes also $s$-channel contributions provides larger predictions at low invariant mass.
For the rapidity difference between the two tagging jets, the hierarchy between the predictions is rather similar.
Therefore, depending on the approximation used, it can vary by $\pm7\%$ and $\pm4\%$ with respect to the full computation at low invariant mass and low rapidity difference for the tagging jets, respectively

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/mjj_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/dyj1j2_NLO.pdf}
\caption{\label{fig:distNLO1} Differential distributions in the invariant mass (left) and rapidity difference of the two tagging jets (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy and order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

Concerning the transverse momentum (left) and rapidity (right) of the hardest jet shown in Fig.~\ref{fig:distNLO2}, the situation is rather different.
While {\sc MadGraph5\_aMC\-@NLO} is very close to the full prediction for low transverse momentum, it departs from it 
at larger transverse momentum by about $10\%$.
This is in contrast with the VBS-approximated predictions such as {\sc Bonsay}, {\sc Powheg}, and {\sc VBFNLO} which are lower than the full computation at low transverse momentum and higher for larger transverse momentum.
The difference at high transverse momentum between the latter predictions and the full computation can be attributed to EW Sudakov logarithms that become large in this phase-space region.
While, the predictions of {\sc Bonsay} and {\sc Powheg} are rather close over the whole range, the one of {\sc VBFNLO} is very different at low transverse momentum where it is even higher than the full computation.
We note that for the transverse momentum of the second hardest jet, the predictions from {\sc MadGraph5\_aMC\-@NLO} are in good agreement with the other VBS-approximated predictions.
% Finally, {\sc VBFNLO} predicts higher rates over the whole range apart from around $200\GeV$ where it is in perfect agreement with the complete calculation.
Concerning the rapidity of the hardest jet, {\sc VBFNLO} is in good agreement with {\sc MoCaNLO}+{\sc Recola} in the rapidity range $|y_{j_1}| < 3$.
For larger rapidity, the other codes constitute a better description of the full process at order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/ptj1_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/yj1_NLO.pdf}
\caption{\label{fig:distNLO2} Differential distributions in the transverse momentum (left) and rapidity of the hardest jet (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy and order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

The last set of differential distributions is the invariant mass of the two charged lepton (left) and the Zeppenfeld variable for the anti-muon (right).
Concerning the comparison of the predictions, both distributions display a rather similar behaviour.
Indeed, the hierarchy mentioned previously is here respected and enhanced towards high invariant mass or high Zeppenfeld variable.
The predictions of {\sc MoCaNLO}+{\sc Recola} and {\sc VBFNLO} are in rather good agreement for both distributions for the kinematic range displayed here.
The other three VBS approximations are close to each other within few per cent.

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/mll_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/zmu_NLO.pdf}
\caption{\label{fig:distNLO3} Differential distributions in the invariant mass of the two charged leptons (left) and Zeppenfeld variable for the muon (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy and order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots presents all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

@@

% As the $s$-channel contributions are sizeable, their interferences with the $t$/$u$-channel can be of similar size in the phase-space region where the former are large.
In phase-space regions where the $s$-channel contributions are sizeable their interference with the $t/u$-channel can be of similar size.
This can be observed by comparing the predictions of {\sc VBFNLO} against the ones of {\sc MadGraph5\_aMC\-@NLO}.

Finally, the effect of EW corrections and non-factorisable contributions in the virtual corrections are usually small.
But they can be relatively large (about $10\%$) for large transverse momentum of the hardest jet.
These high-energy region of the phase space are where EW Sudakov logarithms become large.
Nonetheless these regions are rather suppressed and thus these effects are hardly visible at the level of the cross section.

>>>>

We start with Fig.~\ref{fig:distNLO1} which displays the invariant mass (left) and the rapidity separation (right) of the two tagging jets.
For high invariant mass, all predictions agree rather well.
On the other hand, for low invariant mass, the hierarchy present at the level of the cross section is here reproduced.
The VBS-approximated predictions ({\sc Bonsay} and {\sc Powheg-Box}) are lower than the full calculation ({\sc MoCaNLO}+{\sc Recola}).
The full calculation is rather well approximated by the hybrid VBS approximation implemented in {\sc MG5\_aMC}.
Finally, {\sc VBFNLO} which includes also $s$-channel contributions provides larger predictions at low invariant mass.
For the rapidity difference between the two tagging jets, the hierarchy between the predictions is rather similar.
Therefore, depending on the approximation used, it can vary by $\pm7\%$ and $\pm4\%$ with respect to the full computation at low invariant mass and low rapidity difference for the tagging jets, respectively

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/mjj_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/dyj1j2_NLO.pdf}
\caption{\label{fig:distNLO1} Differential distributions in the invariant mass (left) and rapidity difference of the two tagging jets (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy at $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots present all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

Concerning the transverse momentum (left) and rapidity (right) of the hardest jet shown in Fig.~\ref{fig:distNLO2}, the situation is rather different.
While {\sc MG5\_aMC} is very close to the full prediction for low transverse momentum, it departs from it 
at larger transverse momentum by about $10\%$.
This is in contrast with the VBS-approximated predictions such as {\sc Bonsay}, {\sc Powheg}, and {\sc VBFNLO} which are lower than the full computation at low transverse momentum and higher for larger transverse momentum.
The difference at high transverse momentum between the latter predictions and the full computation can be attributed to EW Sudakov logarithms that become large in this phase-space region.
While, the predictions of {\sc Bonsay} and {\sc Powheg} are rather close over the whole range, the one of {\sc VBFNLO} is very drifferent at low transverse momentum where it is even higher than the full computation.
We note that for the transverse momentum of the second hardest jet, the predictions from {\sc MG5\_aMC} are in good agreement with the other VBS-approximated predictions.
% Finally, {\sc VBFNLO} predicts higher rates over the whole range apart from around $200\GeV$ where it is in perfect agreement with the complete calculation.
Concerning the rapidity of the hardest jet, {\sc VBFNLO} is in good agreement with {\sc MoCaNLO}+{\sc Recola} in the rapidity range $|y_{j_1}| < 3$.
For larger rapidity, the other codes constitute a better description of the full process at order $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/ptj1_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/yj1_NLO.pdf}
\caption{\label{fig:distNLO2} Differential distributions in the transverse momentum (left) and rapidity of the hardest jet (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy at $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots present all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

The last set of differential distributions is the invariant mass of the two charged lepton (left) and the Zeppenfeld variable for the anti-muon (right).
Concerning the comparison of the predictions, both distributions display a rather similar behaviour.
Indeed, the hierarchy mentioned previously is here respected and enhanced towards high invariant mass or high Zeppenfeld variable.
The predictions of {\sc MoCaNLO}+{\sc Recola} and {\sc VBFNLO} are in rather good agreement for both distributions for the kinematic range displayed here.
The other three VBS approximations are close to each other within few per cents.

 \begin{figure*}[hbt!]
   \centering
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/mll_NLO.pdf}
   \includegraphics[width=0.4\textwidth,angle=0,clip=true,trim={0.4cm 2cm 0.cm 1.cm}]{figures/NLO/zmu_NLO.pdf}
\caption{\label{fig:distNLO3} Differential distributions in the invariant mass of the two charged leptons (left) and Zeppenfeld variable for the muon (right).
The LHC process considered is ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ at NLO accuracy at $\mathcal{O}(\alpha_{\rm s}\alpha^6)$.
The description of the different programs used can be found in Sec.~\ref{subsec:codedescr}.
The upper plots provide the absolute value for each prediction while the lower plots present all predictions normalised to {\sc MoCaNLO}+{\sc Recola} which is the full predictions.
The band corresponds to a seven-points scale variation of the renormalisation and factorisation scale.
The predictions are obtained in the fiducial region described in Sec.~\ref{subsec:inputpar}.
}
\end{figure*}

@@

% As the $s$-channel contributions are sizeable, their interferences with the $t$/$u$-channel can be of similar size in the phase-space region where the former are large.
In phase-space regions where the $s$-channel contributions are sizeable their interference with the $t/u$-channel can be of similar size.
This can be observed by comparing the predictions of {\sc VBFNLO} against the ones of {\sc MG5\_aMC}.

Finally, the effect of EW corrections and NF contributions in the virtual corrections are usually small.
But they can be relatively large (about $10\%$) for large transverse momentum of the hardest jet.
These high-energy region of the phase space are where EW Sudakov logarithms become large.
Nonetheless these regions are rather suppressed and thus these effects are hardly visible at the level of the cross section.
